---
title: "Training data analysis"
author: "Leah Kemp"
date: "Date: `r base::format(Sys.time(), '%d/%m/%Y')`"
always_allow_html: true
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: false
    self_contained: yes
  editor_options: 
    chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, fig.align = "center")
```

```{r}
# load packages
library(googlesheets4)
library(tidyverse)
library(janitor)
library(DT)
library(scales)
library(lubridate)
library(plotly)
```

```{r}
# links to excel spreadsheets in google docs
training_plan_july_2021_july_2022 <- "https://docs.google.com/spreadsheets/d/1jSN7xN3CIQUQNUssWz2c9Wz-ab5qCaBYaYvdsNFKSY0/edit#gid=35820724"
training_plan_feb_august_2021 <- "https://docs.google.com/spreadsheets/d/10gONryFN9afkoZvwuJ18rZ_kEO2p_R-H2O85Nr8ktv4/edit#gid=0"

# read in data from google sheets
M3_DAILY_FLEX_EXTRA <- read_sheet(training_plan_july_2021_july_2022,
                                  sheet = "M3_DAILY_FLEX_EXTRA")

M3_POLE_FULL_BODY <- read_sheet(training_plan_july_2021_july_2022,
                                sheet = "M3_POLE_FULL_BODY")

M3_LEG_DAY_EXTRA <- read_sheet(training_plan_july_2021_july_2022,
                               sheet = "M3_LEG_DAY_EXTRA")

M2_FULL_BODY <- read_sheet(training_plan_july_2021_july_2022,
                           sheet = "M2_FULL_BODY")

M2_CORE_FLEX <- read_sheet(training_plan_july_2021_july_2022,
                           sheet = "M2_CORE_FLEX")

M1_POLE <- read_sheet(training_plan_july_2021_july_2022,
                      sheet = "M1_POLE")

M1_GYM <- read_sheet(training_plan_july_2021_july_2022,
                     sheet = "M1_GYM")

targeting_the_requirements <- read_sheet(training_plan_feb_august_2021,
                                         sheet = "targeting_the_requirements")

targeting_the_tricks <- read_sheet(training_plan_feb_august_2021,
                                   sheet = "targeting_the_tricks")

# create a list of the datasets I'm working with
dataset_list <- list(M3_DAILY_FLEX_EXTRA,
                     M3_POLE_FULL_BODY,
                     M3_LEG_DAY_EXTRA,
                     M2_FULL_BODY,
                     M2_CORE_FLEX,
                     M1_POLE,
                     M1_GYM,
                     targeting_the_tricks,
                     targeting_the_requirements)

# setup my function to prepare the weights data
prepare_weights_data <- function(df_list) {
  
  data <- df_list %>%
    # recode empty strings "" by NAs
    na_if("") %>%
    
    # remove empty rows
    janitor::remove_empty(c("rows")) %>%  
    
    # remove some columns to only get the weights data
    # this will also remove columns that don't have a date in the name (data hasn't been entered in those columns)
    dplyr::select(requirement,
                  exercise,
                  matches("\\d{1,2}/\\d{1,2}/\\d{1,4}")) %>%
    select(!contains(c("notes", "reps_or_secs"))) %>%
    
    # format all columns that have a date in their name, to integer format
    mutate_at(vars(matches("\\d{1,2}/\\d{1,2}/\\d{1,4}.set")), as.integer) %>%
    
    # make this data long for downstream analyses
    # this also extracts the date from the column names and makes a variable out of it
    pivot_longer(cols = -c(requirement, exercise),
                 names_pattern = "(\\d{1,2}/\\d{1,2}/\\d{1,4}).(....).(weight)$",
                 names_to = c("date", "set", "weight")) %>%
    mutate(date=ifelse(date=="", "value", date)) %>%
    
    # clean up columns
    select(-weight) %>%
    rename(weight = value) %>%
    
    # remove rows with NA values (exercise not done)
    na.omit()
  
  # format date column as a date
  data$date <- as.Date(data$date,"%d/%m/%y", tz = "NZ")
  
  return(data)
  
}

# setup my function to prepare the reps/secs data
prepare_reps_or_secs_data <- function(df_list) {
  
  data <- df_list %>%
    # recode empty strings "" by NAs
    na_if("") %>%
    
    # remove empty rows
    janitor::remove_empty(c("rows")) %>%  
    
    # remove some columns to only get the sets and reps
    # this will also remove columns that don't have a date in the name (data hasn't been entered in those columns)
    dplyr::select(requirement,
                  exercise,
                  matches("\\d{1,2}/\\d{1,2}/\\d{1,4}")) %>%
    # remove some columns to only get the sets and reps
    select(!contains(c("notes", "weight"))) %>%
    
    # format all columns that have a date in their name, to integer format
    mutate_at(vars(matches("\\d{1,2}/\\d{1,2}/\\d{1,4}.set")), as.integer) %>%
    
    # make this data long for downstream analyses
    # this also extracts the date from the column names and makes a variable out of it
    pivot_longer(cols = -c(requirement, exercise),
                 names_pattern = "(\\d{1,2}/\\d{1,2}/\\d{1,4}).(....).(reps_or_secs)$",
                 names_to = c("date", "set", "reps_or_secs")) %>%
    mutate(date=ifelse(date=="", "value", date)) %>%
    
    # clean up columns
    select(-reps_or_secs) %>%
    rename(reps_or_secs = value) %>%
    
    # remove rows with NA values (exercise not done)
    na.omit()
  
  # format date column as a date
  data$date <- as.Date(data$date,"%d/%m/%y", tz = "NZ")
  
  return(data)
  
}

# apply my function to prepare weights and reps/secs data
weights_data <- lapply(dataset_list, prepare_weights_data)
reps_or_secs_data <- lapply(dataset_list, prepare_reps_or_secs_data)

# collapse dataframes into one dataset
weights_data <- Reduce(rbind, weights_data)
reps_or_secs_data <- Reduce(rbind, reps_or_secs_data)

# merge weights and reps/secs data
data <- full_join(weights_data,
                  reps_or_secs_data,
                  by = c("requirement", "exercise", "date", "set"))

# replace missing values for weight (where I didn't enter a value since no weight was used) with 0
data <- replace(data, is.na(data), 0)

# generalise the requirement list
upper_body_strength <- c("Cup grip handspring deadlift",
                         "Internal and external shoulder rotation",
                         "No leg climb",
                         "Pushing strength",
                         "Shoulder flexion (with external rotation)",
                         "Shoulder flexion (with neutral)",
                         "Shoulder mount deadlift",
                         "True grip ayesha",
                         "Twisted grip handspring deadlift",
                         "Upper body power",
                         "Upper body strength",
                         "Pulling strength")

lower_body_strength <- c("Lower body power",
                         "Lower body strength")

core_strength <- c("Anti-extension core strength",
                   "Anti-lateral core strength",
                   "Anti-rotation core strength")

back_bending_rotation_flexibility <- c("Back flexibility",
                                       "Thoracic rotation",
                                       "Upper back/shoulder flexibility",
                                       "Shoulder flexion flexibility and strength")

splits_flexibility <- c("Active hamstring flexibility",
                        "Active hamstring flexibility/hip flexion strength",
                        "Active hip adduction and abduction",
                        "Front splits",
                        "Hamstring nerve tension",
                        "Hip extension flexibility",
                        "Internal and external hip rotation",
                        "Middle split",
                        "Pancake split",
                        "Point drills",
                        "Microbend drills",
                        "Hip flexor flexibility",
                        "Hamstring flexibility",
                        "Hip flexion strength",
                        "Glute and hamstring strength",
                         "Hip abduction strength",
                         "Hip adduction strength")

# add this broader requirements as a new column in my data
data <- data %>%
  mutate(requirement_broad = case_when(requirement %in% upper_body_strength ~ "Upper body strength",
                                       exercise == "Pole invert starting position isometric hold" ~ "Upper body strength",
                                       exercise == "Pole invert finishing position isometric hold" ~ "Upper body strength",
                                       requirement %in% lower_body_strength ~ "Lower body strength",
                                       requirement %in% core_strength ~ "Core strength",
                                       exercise == "Pole invert knee tucks" ~ "Core strength",
                                       requirement %in% back_bending_rotation_flexibility  ~ "Back bending and rotation flexibility",
                                       requirement %in% splits_flexibility ~ "Splits flexibility"))

# setup plotting functions
time_series_plot <- function(data, wrap_variable) {
  data %>%
    ggplot(aes(x = date, y = number_of_sets_completed)) +
    geom_bar(stat="identity") +
    theme(legend.position = "none") +
    facet_wrap(vars(exercise), ncol = 1) +
    geom_vline(xintercept = as.numeric("18/12/2021"), colour = "red") +
    labs(title = "",
         x = "Date",
         y = "Number of sets") +
    theme_bw() +
    theme(axis.text.x = element_text(colour = "grey20", size = 8, angle = 90,
                                     hjust = 0.5, vjust = 0.5),
          axis.text.y = element_text(colour = "grey20", size = 8),
          legend.title = element_blank(),
          text = element_text(size = 8)) +
    # create a vertical line in plots at certain date where all my training data was collected (finished up with Jhene)
    scale_x_date(labels = date_format("%Y-%m-%d"),
                 breaks = "1 month", date_labels = paste0("%m", "/", "%Y"))
}

time_series_summary_broad_plot <- function(data) {
  data %>%
    ggplot(aes(x = date, y = number_of_sets_completed)) +
    geom_bar(stat="identity") +
    theme(legend.position = "none") +
    facet_wrap(vars(requirement_broad), ncol = 1) +
    geom_vline(xintercept = as.numeric("18/12/2021"), colour = "red") +
    labs(title = "",
         x = "Date",
         y = "Number of sets") +
    theme_bw() +
    theme(axis.text.x = element_text(colour = "grey20", size = 8, angle = 90,
                                     hjust = 0.5, vjust = 0.5),
          axis.text.y = element_text(colour = "grey20", size = 8),
          legend.title = element_blank(),
          text = element_text(size = 8)) +
    # create a vertical line in plots at certain date where all my training data was collected (finished up with Jhene)
    scale_x_date(labels = date_format("%Y-%m-%d"),
                 breaks = "1 month", date_labels = paste0("%m", "/", "%Y"))
}

time_series_summary_plot <- function(data) {
  data %>%
    ggplot(aes(x = date, y = number_of_sets_completed)) +
    geom_bar(stat="identity") +
    theme(legend.position = "none") +
    facet_wrap(vars(requirement), ncol = 1) +
    geom_vline(xintercept = as.numeric("18/12/2021"), colour = "red") +
    labs(title = "",
         x = "Date",
         y = "Number of sets") +
    theme_bw() +
    theme(axis.text.x = element_text(colour = "grey20", size = 8, angle = 90,
                                     hjust = 0.5, vjust = 0.5),
          axis.text.y = element_text(colour = "grey20", size = 8),
          legend.title = element_blank(),
          text = element_text(size = 8)) +
    # create a vertical line in plots at certain date where all my training data was collected (finished up with Jhene)
    scale_x_date(labels = date_format("%Y-%m-%d"),
                 breaks = "1 month", date_labels = paste0("%m", "/", "%Y"))
}
```

## PR

```{r, out.width = "100%"}
# get the combinations of weights and reps/sec for each exercise
weight_reps_pairs <- data %>%
  select(-c(set, requirement, requirement_broad)) %>%
  group_by(exercise, weight, reps_or_secs) %>%
  summarise(weight, reps_or_secs)

# split dataset by exercise
weight_reps_pairs_by_exercise <- split(weight_reps_pairs, weight_reps_pairs$exercise)

get_pr <- function(data_list) {
  
  # order data by highest weight and highest number of reps/secs (priority to highest weight)
  ordered <- data_list %>%
    arrange(desc(weight), desc(reps_or_secs))
  
  # get the top row
  pr <- ordered[1, ]
  
  return(pr)
  
}

# apply my function to my list of datasets (so the pr is calculated for each exercise)
pr <- lapply(weight_reps_pairs_by_exercise, get_pr)

# collapse results into a single dataframe
pr <- Reduce(rbind, pr)

# put in table
pr %>%
  DT::datatable(filter = "top",
                colnames = c("Exercise",
                             "Max weight",
                             "Reps/secs"),
                rownames = FALSE,
                extensions = list("ColReorder" = NULL,
                                  "Buttons" = NULL,
                                  "FixedColumns" = list(leftColumns=1)),
                options = list(
                  dom = "BRrltpi",
                  autoWidth = TRUE,
                  buttons =
                    list("copy",
                         list(extend = "collection",
                              buttons = c("csv", "excel", "pdf"),
                              text = "Download"),
                         I("colvis"))))
```

## Excercise list

```{r, out.width = "100%"}
# create table of list of exercises
data %>%
  select(exercise,
         requirement_broad) %>%
  unique() %>%
  # order alphabetically by requirement
  arrange(exercise) %>%
  # make factors for table formatting
  mutate(requirement_broad = as.factor(requirement_broad)) %>%
  DT::datatable(filter = "top",
                colnames = c("Exercise",
                             "Requirement - broad"),
                rownames = FALSE,
                extensions = list("ColReorder" = NULL,
                                  "Buttons" = NULL,
                                  "FixedColumns" = list(leftColumns=1)),
                options = list(
                  dom = "BRrltpi",
                  autoWidth = TRUE,
                  buttons =
                    list("copy",
                         list(extend = "collection",
                              buttons = c("csv", "excel", "pdf"),
                              text = "Download"),
                         I("colvis"))))
```

## Sets completed - time series {.tabset .tabset-fade}

*Note. no lower body/upper body and core strength data available from personal training sessions between Dec 2020 to Dec 2021*

```{r}
# summarise data by broad requirement and date
time_series_data <- data %>%
  group_by(requirement_broad, date) %>%
  summarise(number_of_sets_completed = n())
```

### All categories

```{r, out.width = "100%"}
# split data by broad requirement
time_series_data <- time_series_data %>% split(time_series_data$requirement_broad)

# plot
plot_ly(type = "scatter",
        mode = "lines",
        texttemplate="%{label}") %>%
  add_trace(data = time_series_data$`Back bending and rotation flexibility`,
            x = ~date,
            y = ~number_of_sets_completed,
            name = "Back bending and rotation flexibility") %>%
  add_trace(data = time_series_data$`Core strength`,
            x = ~date,
            y = ~number_of_sets_completed,
            name = "Core strength") %>%
  add_trace(data = time_series_data$`Lower body strength`,
            x = ~date,
            y = ~number_of_sets_completed,
            name = "Lower body strength") %>%
  add_trace(data = time_series_data$`Splits flexibility`,
            x = ~date,
            y = ~number_of_sets_completed,
            name = "Splits flexibility") %>%
  add_trace(data = time_series_data$`Upper body strength`,
            x = ~date,
            y = ~number_of_sets_completed,
            name = "Upper body strength") %>%
  layout(showlegend = TRUE,
         xaxis = list(rangeslider = list(visible = TRUE),
                      rangeselector = list(
                        buttons = list(
                          list(count=1, label="1m", step="month", stepmode="backward"),
                          list(count=6, label="6m", step="month", stepmode="backward"),
                          list(count=1, label="YTD", step="year", stepmode="todate"),
                          list(count=1, label="1y", step="year", stepmode="backward"),
                          list(step="all"))))) %>%
  layout(xaxis = list(title = "Date",
                      zerolinecolor = "#ffff",
                      zerolinewidth = 2,
                      gridcolor = "ffff"),
         yaxis = list(title = "Number of sets completed",
                      zerolinecolor = "#ffff",
                      zerolinewidth = 2,
                      gridcolor = "ffff"),
         plot_bgcolor="#e5ecf6", width = 900)
```

### Core strength

```{r}
# summarise data by broad requirement, exercise and date
time_series_data <- data %>%
  group_by(requirement_broad, exercise, date) %>%
  summarise(number_of_sets_completed = n())
```

```{r, fig.height = 35, out.width = "100%"}
time_series_data %>%
  filter(requirement_broad == "Core strength") %>%
  time_series_plot()
```

### Lower body strength

```{r, fig.height = 20, out.width = "100%"}
time_series_data %>%
  filter(requirement_broad == "Lower body strength") %>%
  time_series_plot()
```

### Upper body strength

```{r, fig.height = 35, out.width = "100%"}
time_series_data %>%
  filter(requirement_broad == "Upper body strength") %>%
  time_series_plot()
```

### Splits flexibility

```{r, fig.height = 40, out.width = "100%"}
time_series_data %>%
  filter(requirement_broad == "Splits flexibility") %>%
  time_series_plot()
```

### Back bending and rotation flexibility

```{r, fig.height = 30, out.width = "100%"}
time_series_data %>%
  filter(requirement_broad == "Back bending and rotation flexibility") %>%
  time_series_plot()
```

## Stretching

### Weekly total time stretching

**Target: 5-10 minutes per week per muscle**

```{r, out.width = "100%"}
# recode empty strings "" by NAs
M3_DAILY_FLEX_EXTRA <- M3_DAILY_FLEX_EXTRA %>%
  na_if("") 

# format all columns that have a date in their name, to integer format
M3_DAILY_FLEX_EXTRA <- M3_DAILY_FLEX_EXTRA %>%
  mutate_at(vars(matches("\\d{1,2}/\\d{1,2}/\\d{1,2}.set")), as.integer)

# remove some columns to only get the sets and reps
# this'll also remove columns that don't have a date in the name (data hasn't been entered in those columns)
M3_DAILY_FLEX_EXTRA <- M3_DAILY_FLEX_EXTRA %>%
  dplyr::select(requirement,
                exercise,
                matches("\\d{1,2}/\\d{1,2}/\\d{1,2}"))

# make this data long for downstream analyses
# this also extracts the date from the column names and makes a variable out of it
M3_DAILY_FLEX_EXTRA_long <- pivot_longer(M3_DAILY_FLEX_EXTRA,
                       cols = -c(requirement, exercise),
                       names_pattern = "(\\d{1,2}/\\d{1,2}/\\d{1,2}).(....).(reps_or_secs)$",
                       names_to = c("date", "set", "reps_or_secs")) %>% 
  mutate(date=ifelse(date=="", "value", date))

# clean up columns
M3_DAILY_FLEX_EXTRA_long <- M3_DAILY_FLEX_EXTRA_long %>%
  select(-reps_or_secs) %>%
  rename(reps_or_secs = value)

# remove rows with NA values (exercise not done)
M3_DAILY_FLEX_EXTRA_long <- M3_DAILY_FLEX_EXTRA_long %>%
  na.omit()

# format date column as a date
M3_DAILY_FLEX_EXTRA_long$date <- as.Date(M3_DAILY_FLEX_EXTRA_long$date,"%d/%m/%y")

# create a column that specifies the week
M3_DAILY_FLEX_EXTRA_long <- M3_DAILY_FLEX_EXTRA_long %>%
  dplyr::mutate(week = lubridate::week(lubridate::ymd(M3_DAILY_FLEX_EXTRA_long$date)))
```

```{r}
# summarise data by week and exercise
time_series_data <- M3_DAILY_FLEX_EXTRA_long %>%
  group_by(requirement, exercise, week) %>%
  summarise(number_of_reps_or_secs_completed = sum(reps_or_secs))
```

```{r, out.width = "100%"}
# a function for getting integer y-axis values
# (grabbed code from: https://joshuacook.netlify.app/post/integer-values-ggplot-axis/)
integer_breaks <- function(n = 5, ...) {
  fxn <- function(x) {
    breaks <- floor(pretty(x, n, ...))
    names(breaks) <- attr(breaks, "labels")
    breaks
  }
  return(fxn)
}

time_series_data %>%
  ggplot(aes(x = week, y = number_of_reps_or_secs_completed)) +
  geom_bar(stat="identity") +
  theme(legend.position = "none") +
  facet_wrap(vars(requirement)) +
  labs(title = "",
       x = "Week",
       y = "Total seconds") +
  theme_bw() +
  theme(axis.text.x = element_text(colour = "grey20", size = 8, angle = 90,
                                   hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(colour = "grey20", size = 8),
        legend.title = element_blank(),
        text = element_text(size = 8)) +
  geom_hline(yintercept = 300, colour = "orange") +
  geom_hline(yintercept = 600, colour = "forest green") +
  scale_x_continuous(breaks = integer_breaks())
```

### Stretching frequency

**Target: 5-6 days week per week**

```{r}
# summarise data by week and exercise
time_series_data <- M3_DAILY_FLEX_EXTRA_long %>%
  group_by(requirement, week) %>%
  summarise(number_of_sessions_completed = length(unique(date)))
```

```{r, out.width = "100%"}
time_series_data %>%
  ggplot(aes(x = week, y = number_of_sessions_completed)) +
  geom_bar(stat="identity") +
  theme(legend.position = "none") +
  facet_wrap(vars(requirement)) +
  labs(title = "",
       x = "Week",
       y = "Total number of sessions") +
  theme_bw() +
  theme(axis.text.x = element_text(colour = "grey20", size = 8, angle = 90,
                                   hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(colour = "grey20", size = 8),
        legend.title = element_blank(),
        text = element_text(size = 8)) +
  geom_hline(yintercept = 5, colour = "orange") +
  geom_hline(yintercept = 6, colour = "forest green") +
  scale_x_continuous(breaks = integer_breaks())
```

```{r, cleanup, results = "hide"}
# clean up
rm(list = ls())
```
